---
title: "DYB Summary Heatmaps"
author: "Shabrina Mardevi"
output:
  pdf_document: default
  html_document: default
---

```{r error=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(dplyr)
library(janitor)
library(tidyverse)
library(ggplot2)
library(readxl)
library(openxlsx)
library(data.table)
library(pointblank)
library(haven)
library(naniar)
library(pointblank)
library(heatmaply)
library(ggpubr)
```

# Summary Heatmaps: Data Availability (All Categories) by Number of Years from 2009 to 2018 (Blue Heatmaps)

## Prep Data
```{r}
# Create function to :1.) Join two 5-year period data, and 2.) Add the n years of availability

join_transform <- function(table_1, table_2){
  table_all <- full_join(table_1, table_2, by = c("countries_areas", "data_available"))
  table_clean <-
    table_all %>% 
    mutate_at(vars(n.x, n.y), ~replace(., is.na(.), 0)) %>%  
    mutate (n = n.x + n.y) %>% 
    select(-c(n.x, n.y))
  
  return(table_clean)
}
``` 

### DYB Table 4 - Vital statistics summary and life expectancy at birth:  2014 - 2018
### Live Births, Infant Deaths, and Deaths (General) by Countries

```{r}

# Create Function to Transform Table 4: Number of Years Available for 
#Live Births, Infant Deaths, and Deaths (General) by Countries
t4_transform <- function(table, years, table_id){

# Prep original
table_clean <-
  table[- 1, ]   %>%
  select (c("...1", contains("number"))) %>% 
  rename_all(~c("countries_areas",
                "live_births",
                "deaths",
                "infant_deaths")) %>% 
  mutate(year=countries_areas) %>% 
  replace_with_na(replace = list(countries_areas = years)) %>% 
  separate(countries_areas, c("countries_areas"), " - ") %>% 
  fill(countries_areas) %>% 
  replace_with_na_all(condition = ~.x == "...") %>% 
  replace_with_na_all(condition = ~.x == "-") 

# Create count dataframe
table_count <- 
  table_clean %>%
  gather(key="data_available", value="value", -c(countries_areas, year)) %>% 
  mutate (countries_areas = gsub('[0-9]+', '', countries_areas)) %>%
  filter(value != is.na(.)) %>% 
  group_by(countries_areas, data_available) %>% 
  summarise(n=n())

return(table_count)
}
```

```{r error=FALSE, warning=FALSE, message=FALSE}
# Load Table 4s
table4_DYB_all_2 <- read_excel("table04.xlsx", skip = 4) #2014-2018
table4_DYB_all_1 <- read_excel("Table04 (1).xlsx", skip = 4) #2009-2013
```

```{r error=FALSE, warning=FALSE, message=FALSE}
# Transform Table 4s
table4_1 <- t4_transform(table4_DYB_all_1, 
                         c("2009", "2010", "2011", "2012", "2013"))
table4_2 <- t4_transform(table4_DYB_all_2, 
                         c("2014", "2015", "2016", "2017", "2018"))
``` 


```{r error=FALSE, warning=FALSE, message=FALSE}

# Merge Table 4s :2009-2018
table4_clean <- join_transform(table4_1, table4_2)

``` 

### DYB Table 22 - Marriages and crude marriage rates, by urban/rural residence:  2014 - 2018
### DYB Table 24 - Divorces and crude divorce rates by urban/rural residence:  2014 - 2018
```{r}
# Create Function to Transform Table 22-25: Number of Years Available for
# Marriage and Divorces by Countries
ur_transform <- function(data_all, data_type) {

data <-
  data_all [-1,]   %>%
  select (c("...1", contains("20"))) %>% 
  select (1:6) %>% 
  rename_at(.vars = 1, ~c("countries_areas")) %>% 
  separate(countries_areas, c("countries_areas"), " - ") %>% 
  filter (!(countries_areas %in% c("Urban", "Rural"))) %>% 
  replace_with_na(replace = list(countries_areas = "Total")) %>% 
  fill(countries_areas) %>% 
  replace_with_na_all(condition = ~.x == "...") %>% 
  mutate (countries_areas = gsub('[0-9]+', '', countries_areas)) 

data_final <- 
  data %>%
  gather(key="year", value=value, -c(countries_areas)) %>%
  filter(value != is.na(.)) %>% 
  mutate(data_available = data_type) %>% 
  group_by(countries_areas, data_available) %>% 
  summarise(n=n())

return(data_final)

}

```

```{r error=FALSE, warning=FALSE, message=FALSE}
#Load Tables 22-25
table22_DYB <- read_excel("table22.xlsx", skip = 4) #2014-2018
table23_DYB <- read_excel("table23_2013.xlsx", skip = 4) #2009-2013
table24_DYB <- read_excel("table24.xlsx", skip = 4) #2014-2018
table25_DYB <- read_excel("Table25.xlsx", skip = 4) #2009-2013
```

```{r error=FALSE, warning=FALSE, message=FALSE}
#Transform Tables 22-25
table_m2 <- ur_transform(table22_DYB, "marriages")
table_m1 <- ur_transform(table23_DYB, "marriages")
table_d2 <- ur_transform(table24_DYB, "divorces")
table_d1 <- ur_transform(table25_DYB, "divorces")
```

```{r error=FALSE, warning=FALSE, message=FALSE}
table_mall <- join_transform(table_m1, table_m2)
table_dall <- join_transform(table_d1, table_d2)
``` 

# Foetal Deaths by Countries
```{r error=FALSE, warning=FALSE, message=FALSE}
# Load Table 12s
table12_DYB_all_2 <- read_excel("./Gen Death/table12_2.xlsx", skip = 4) #2014-2018
table12_DYB_all_1 <- read_excel("./Gen Death/table12_1.xlsx", skip = 4) #2009-2013
```

```{r error=FALSE, warning=FALSE, message=FALSE}
# Transform Table 4s
table12_1 <- ur_transform(table12_DYB_all_1, "foetal_deaths")
table12_2 <- ur_transform(table12_DYB_all_2, "foetal_deaths")
``` 

```{r error=FALSE, warning=FALSE, message=FALSE}

# Merge Table 4s :2009-2018
table12_clean <- join_transform(table12_1, table12_2)

``` 

## Merge lookup with DYB data
```{r error=FALSE, warning=FALSE, message=FALSE}
# Load UNPA Countries + Regions Lookup
lookup_table_all <- 
  as_tibble(read_excel("UNFPA_countries.xlsx"))

all_tables_dyb <- rbind(table_mall, table_dall, table4_clean, table12_clean)
``` 

### Create lookup consisting all Indicators (5 per country) and Scores

```{r}
data_available <- c("live_births", "infant_deaths", "marriages", "divorces", "deaths", "foetal_deaths")

lookup_score<- function(avail, data) {

# Compute completeness Score by Category
complete_score <-
  data %>% 
    group_by(countries_areas) %>% 
    summarise(complete = n())

# Compute completeness Score by Overall Sum by Country
count_score <-
  aggregate(data$n, by = list(data$countries_areas), sum)
count_score <-
  rename(count_score, "countries_areas" = Group.1)

# Join lookup and indicators
available_count_lookup <- left_join(merge(avail, lookup_table_all),
                                   complete_score, by="countries_areas")
available_count_lookup <-
  rename(available_count_lookup, "data_available" = x)

available_count_lookup <- left_join(available_count_lookup, count_score, by = "countries_areas")

}
```

```{r merge data with lookup}

data_available_lookup <- lookup_score(data_available, all_tables_dyb)

#Create a UNFPA Life Course Approach Complete Data
UNFPA_lifeapp <-
  left_join(data_available_lookup, all_tables_dyb, 
            by = c("countries_areas", "data_available"), ignore_case = T)

UNFPA_lifeapp <-
  UNFPA_lifeapp %>% 
  mutate_all(~replace(., is.na(.), 0)) %>% 
  # Create score :0.75 on completeness across categories + 0.25 completeness in years
  mutate(score = ((complete/5)*0.85) # scored by completeness across categories 
         + ((x/50)*0.15)) %>% # scored by completeness across years
  mutate (countries_areas = reorder(countries_areas, score))

head(data_available_lookup, 5)
```

## Generate Summary Completeness Heatmaps by Region
```{r fig.height = 6.5, fig.width = 10}

# Set parameters for levels and labels
levels_pref <- c("foetal_deaths", "live_births", "infant_deaths", "marriages", "divorces", "deaths")
labels_pref <- c("Foetal Deaths", "Live Births", "Infant Deaths", "Marriages", "Divorces", "Deaths")
```

```{r fig.height = 6.5, fig.width = 10}

# Create function to produce Heatmaps
num_years_plot <- function(region)
  ggplot(subset(UNFPA_lifeapp, UNFPA_Regions %in% region), 
         aes(x = factor(data_available, 
                        levels = levels_pref), y = countries_areas, fill = n)) + 
  geom_tile() + 
  scale_fill_distiller(name = "Number of\nYear(s) Available", palette = "Blues", 
                       direction = +1, breaks = c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10),
                       limits=c(0,10)) +
  theme_minimal() +
  labs(title = paste("Available Data on Registered Vital Events in",
                     ifelse(region == "AP", "Asia and the Pacific",
                               ifelse(region == "AS", "Arab States",
                                      ifelse(region == "EECA", "Eastern Europe and Central Asia",
                                             ifelse(region == "ESA", "Eastern and Southern Africa",
                                                    ifelse(region == "LAC", 
                                                           "Latin America and Caribbean",
                                                           ifelse(region == "WCA",
                                                                  "Western and Central Africa")))))), 
                     "Region",
                     "\nby Country and Vital Event(s), 2009-2018",
                     sep = " "),
       x = "Registered Vital Events Data", y = "Countries/Areas") +
  scale_x_discrete(labels = labels_pref)
```

```{r fig.height = 8, fig.width = 12}
# Run heatmaps
num_years_plot("AS")
num_years_plot("AP")
num_years_plot("EECA")
num_years_plot("ESA")
num_years_plot("LAC")
num_years_plot("WCA")
```


# Summary Heatmaps: Data Availability Year from 2009 to 2018 (Purple Heatmaps)

This visualization will be remodelled with the following prompts:
- Reordered by countries (most complete to least, top-bottom)
- Facets removed

## Prepare Data
```{r}
# Create table 4 transformation function that includes years and recodes values to binary
# (1 = "Available", 0 = "Unavailable)
t4_year_transform <- function(t4_table, years)
  t4_table[- 1, ]   %>%
  select (c("...1", contains("number"))) %>% 
  rename_all(~c("countries_areas", "live_births", "deaths", "infant_deaths")) %>% 
  mutate(year=countries_areas) %>% 
  replace_with_na(replace = list(countries_areas = years)) %>% 
  separate(countries_areas, c("countries_areas"), " - ") %>% 
  fill(countries_areas) %>% 
  replace_with_na_all(condition = ~.x == "...") %>% 
  replace_with_na_all(condition = ~.x == "-")   %>% 
  filter(!(live_births %in% NA & deaths %in% NA & infant_deaths %in% NA)) %>%
  mutate_at(vars(live_births, deaths, infant_deaths), ~replace(., is.na(.), 0)) %>%
  mutate(live_births = ifelse(live_births > 0, 1, 0)) %>% 
  mutate(deaths = ifelse(deaths > 0, 1, 0)) %>% 
  mutate(infant_deaths = ifelse(infant_deaths > 0, 1, 0)) %>% 
  gather(-c(countries_areas, year), key = "data_available", value = "availability") %>% 
  mutate (countries_areas = gsub('[0-9]+', '', countries_areas))
```

```{r error=FALSE, warning=FALSE, message=FALSE}
# Transform table 4s
table4_1_year <- 
  t4_year_transform(table4_DYB_all_1, 
                    c("2009", "2010", "2011", "2012", "2013"))
table4_2_year <- 
  t4_year_transform(table4_DYB_all_2,
                    c("2014", "2015", "2016", "2017", "2018")) 
```

```{r}

# Create table 22-25 transformation function that includes years and recodes values to binary 
# (1 = "Available", 0 = "Unavailable)
ur_year_transform <- function(data_ur, names, data_type){
  data_ur[- 1, ]   %>%
  select (c("...1", contains("20"))) %>% 
  select (1:6) %>% 
  rename_all(~names) %>% 
  separate(countries_areas, c("countries_areas"), " - ") %>% 
  filter (!(countries_areas %in% c("Urban", "Rural"))) %>% 
  replace_with_na(replace = list(countries_areas = "Total")) %>% 
  fill(countries_areas) %>% 
  replace_with_na_all(condition = ~.x == "...") %>% 
  replace_with_na_all(condition = ~.x == "-") %>% 
  mutate (countries_areas = gsub('[0-9]+', '', countries_areas)) %>%
  gather(-c(countries_areas), key = "year", value = "availability") %>% 
  mutate_at(vars(availability), ~replace(., is.na(.), 0)) %>%
  mutate(availability = ifelse(availability > 0, 1, 0)) %>% 
  mutate(data_available = data_type) %>% 
  select(countries_areas, year, data_available,availability)
  }

```

```{r error=FALSE, warning=FALSE, message=FALSE}
# Transform 22-25, Marriages and Divorces
table_m2_year <- 
  ur_year_transform(table22_DYB, 
                    c("countries_areas", "2014", "2015", "2016", "2017", "2018"), "marriages")
table_m1_year <- 
  ur_year_transform(table23_DYB, 
                    c("countries_areas", "2009", "2010", "2011", "2012", "2013"), "marriages")
table_d2_year <- 
  ur_year_transform(table24_DYB, 
                    c("countries_areas", "2014", "2015", "2016", "2017", "2018"), "divorces")
table_d1_year <- 
  ur_year_transform(table25_DYB,
                    c("countries_areas", "2009", "2010", "2011", "2012", "2013"), "divorces")
```

```{r error=FALSE, warning=FALSE, message=FALSE}
# Transform 12, Foetal Deaths
table_f2_year <- 
  ur_year_transform(table12_DYB_all_2, 
                    c("countries_areas", "2014", "2015", "2016", "2017", "2018"), "foetal_deaths")
table_f1_year <- 
  ur_year_transform(table12_DYB_all_1, 
                    c("countries_areas", "2009", "2010", "2011", "2012", "2013"), "foetal_deaths")
```

```{r error=FALSE, warning=FALSE, message=FALSE}
# Bind all Data
all_tables_year <- rbind(table4_1_year, table4_2_year, table_d1_year, 
                         table_d2_year, table_m1_year, table_m2_year,
                         table_f2_year, table_f1_year)
```

### Create lookup consisting all Indicators (5 per country) and Year
```{r error=FALSE, warning=FALSE, message=FALSE}
# Create Lookup  :Country, Region, Available Data, Year
data_year_lookup <-
  data_available_lookup %>% # Reuse previous lookup that already includes Available Data
  select(data_available:UNFPA_Regions) %>% 
  merge(., c("2009", "2010", "2011", "2012", "2013", "2014", "2015", "2016", "2017", "2018")) %>% 
  # Merge available numbers of years possible
  rename("year" = y)
```

### Generate Data
```{r error=FALSE, warning=FALSE, message=FALSE}
# Merge lookup
table_year_unfpa <- 
  data_year_lookup %>% 
  left_join(., all_tables_year, by=c("countries_areas", "data_available", "year")) %>% 
  mutate_at(vars(availability), ~replace(., is.na(.), 0)) %>% 
  mutate(countries_areas = reorder(countries_areas, availability))
```

## Generate Summary Yearly Availability Heatmaps by Data Category
```{r fig.height = 14, fig.width = 10}

all_plot_sum <- function(region, data_type) {
ggplot(subset(table_year_unfpa, c((data_available %in% data_type)&(UNFPA_Regions %in% region))), aes(x = year, y = countries_areas, fill= availability, color = "grey")) + 
  geom_tile() + 
  scale_fill_distiller(palette = "Purples", direction = +1, breaks = c(0, 1), limits=c(0,1)) +
  theme(axis.text.x = element_text(size =0.1)) +
  theme_classic()+
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank()) +
  theme(legend.position = "none")
}
```

###### Figured out a way to combine separate heatmaps, but have yet to figure out how to align them properly

```{r fig.height = 5, fig.width = 10}
ap_lb <- all_plot_sum("AP", "live_births")
as_lb <- all_plot_sum("AS", "live_births")
eeca_lb <- all_plot_sum("EECA", "live_births")
esa_lb <- all_plot_sum("ESA", "live_births")
lac_lb <- all_plot_sum("LAC", "live_births")
wca_lb <- all_plot_sum("WCA", "live_births")
```

```{r fig.height = 17, fig.width = 10}
ggarrange(eeca_lb, lac_lb, as_lb, ap_lb, esa_lb, wca_lb, 
          labels = c("EECA", "LAC", "AS", "AP", "ESA", "WCA"),
          heights = c(2.2, 4.8, 2, 4.2, 3, 3),
          ncol = 1, nrow = 6)
```

```{r}
ap_id <- all_plot_sum("AP", "infant_deaths")
as_id <- all_plot_sum("AS", "infant_deaths")
eeca_id <- all_plot_sum("EECA", "infant_deaths")
esa_id <- all_plot_sum("ESA", "infant_deaths")
lac_id <- all_plot_sum("LAC", "infant_deaths")
wca_id <- all_plot_sum("WCA", "infant_deaths")
```

```{r fig.height = 17, fig.width = 10}
ggarrange(eeca_id, lac_id, as_id, ap_id, esa_id, wca_id, 
          labels = c("EECA", "LAC", "AS", "AP", "ESA", "WCA"),
          heights = c(2.2, 4.8, 2, 4.2, 3, 3),
          ncol = 1, nrow = 6)
```

```{r fig.height = 5, fig.width = 10}
eeca_lb <- all_plot_sum("EECA", "live_births")
all_plot_sum("EECA", "infant_deaths")
all_plot_sum("EECA", "marriages")
all_plot_sum("EECA", "divorces")
all_plot_sum("EECA", "deaths")
all_plot_sum("EECA", "foetal_deaths")
```

```{r fig.height = 5, fig.width = 10}
ap_id <- all_plot_sum("AP", "infant_deaths")
as_id <- all_plot_sum("AS", "infant_deaths")
eeca_id <- all_plot_sum("EECA", "infant_deaths")
esa_id <- all_plot_sum("ESA", "infant_deaths")
lac_id <- all_plot_sum("LAC", "infant_deaths")
wca_id <- all_plot_sum("WCA", "infant_deaths")
```

```{r fig.height = 17, fig.width = 10}
ggarrange(eeca_id, lac_id, as_id, ap_id, esa_id, wca_id, 
          labels = c("EECA", "LAC", "AS", "AP", "ESA", "WCA"),
          heights = c(2.2, 4.8, 2, 4.2, 3, 3),
          ncol = 1, nrow = 6)
```

```{r fig.height = 5, fig.width = 10}
esa_lb <- all_plot_sum("ESA", "live_births")
all_plot_sum("ESA", "infant_deaths")
all_plot_sum("ESA", "marriages")
all_plot_sum("ESA", "divorces")
all_plot_sum("ESA", "deaths")
all_plot_sum("ESA", "foetal_deaths")
```

```{r fig.height = 5, fig.width = 10}
lac_lb <- all_plot_sum("LAC", "live_births")
all_plot_sum("LAC", "infant_deaths")
all_plot_sum("LAC", "marriages")
all_plot_sum("LAC", "divorces")
all_plot_sum("LAC", "deaths")
all_plot_sum("LAC", "foetal_deaths")
```

```{r fig.height = 5, fig.width = 10}
wca_lb <- all_plot_sum("WCA", "live_births")
all_plot_sum("WCA", "infant_deaths")
all_plot_sum("WCA", "marriages")
all_plot_sum("WCA", "divorces")
all_plot_sum("WCA", "deaths")
all_plot_sum("WCA", "foetal_deaths")
```

# Fac

```{r fig.height = 14, fig.width = 10}

all_plot_sum_facet <- function(data_type) {
ggplot(subset(table_year_unfpa, data_available %in% data_type), 
       aes(x = year, y = countries_areas, fill= availability, color = "grey")) + 
  geom_tile() + 
  facet_grid(UNFPA_Regions~., space= "free_y", scales = "free_y")+
  scale_fill_distiller(palette = "Purples", direction = +1, breaks = c(0, 1), limits=c(0,1)) +
  theme(axis.text.x = element_text(size =0.1)) +
  theme_classic()+
  theme(legend.position = "none")+
  labs(title = paste("Available Data of Registered", 
                     ifelse(data_type=="live_births", "Live Births",
                            ifelse(data_type == "infant_deaths", "Infant Deaths",
                                   ifelse(data_type == "marriages", "Marriages",
                                          ifelse(data_type=="divorces", "Divorces",
                                                 ifelse(data_type=="deaths", "General Deaths",
                                                        ifelse(data_type == "foetal_deaths", "Foetal Deaths")))))),
                     "Region, 2009-2018"), 
       x = "Year", y = "Country/Area") 
}
```

```{r fig.height = 12, fig.width = 10}
all_plot_sum_facet("live_births")
all_plot_sum_facet("infant_deaths")
all_plot_sum_facet("marriages")
all_plot_sum_facet("divorces")
all_plot_sum_facet("deaths")
all_plot_sum_facet("foetal_deaths")
```




# Detailed Heatmaps: Numbers of Years Available by Types of Registration from 2009 to 2018 (Green Heatmaps)

```{r error=FALSE, warning=FALSE, message=FALSE}

# Create Function to Generate Detailed (Green) Heatmaps

det_heatmap <- function(data, region, title, labels, levels) {

ggplot(subset(data, UNFPA_Regions %in% region), aes(x = factor(data_available, levels = levels), y = countries_areas, fill= n)) + 
  geom_tile() + 
  scale_fill_distiller(name = "Number of \n Year(s) Available", palette = "Greens", 
                       direction = +1, breaks = c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10),
                       limits=c(0,10)) +
  theme_minimal() +
  labs(x = paste(title, "Data", sep = " "), y = "Countries/Areas") +
  labs(title = paste("Available Data on Registered",
                     title,
                     "in the",
                     ifelse(region == "AP", "Asia and the Pacific",
                               ifelse(region == "AS", "Arab States",
                                      ifelse(region == "EECA", "Eastern Europe and Central Asia",
                                             ifelse(region == "ESA", "Eastern and Southern Africa",
                                                    ifelse(region == "LAC", "Latin America and Caribbean",
                                                           ifelse(region == "WCA", "Western and Central Africa")))))), 
                     "Region", 
                      "\nby Country and Disaggregation Variable(s), 2009-2018",
                     sep = " "),
       x = paste(title, "Data"), y = "Countries/Areas") +
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  scale_x_discrete(labels = labels)
}

```

## Live Births
```{r error=FALSE, warning=FALSE, message=FALSE}

# Table Live Births

all_bt <- list.files(path = "./Birth Regs", 
                        pattern = "table",
                        full.names = TRUE)

all_bts_excel <-
  lapply(all_bt, function (i) {
  x = read_excel(i, skip = 4)
}) 
```

```{r error=FALSE, warning=FALSE, message=FALSE}

# Create Freq Table Function for non-Urban/Rural DYB Tables
ap_transform <- function(data, data_type) {
  data[- 1, ] %>%
  dplyr::select (1:2) %>% 
  rename_all(~c("countries_areas", "total")) %>% 
  mutate(countries_areas = gsub('[0-9]+', '', countries_areas)) %>% 
  filter (!(countries_areas %in% c(" - ", " +", " (C)", " (+U)", " (|)", " (U)", 
                                   " (+C)", "* (C)", "* (+C)", "* (+U)", 
                                   "Unknown - Inconnu", "* (U)"))) %>% 
  replace_with_na(replace = list(countries_areas = "Total")) %>% 
  fill(countries_areas) %>% 
  filter(!(total == is.na(.))) %>% 
  separate(countries_areas, c("countries_areas"), " - ") %>% 
  group_by(countries_areas) %>% 
  summarise(n=n()) %>% 
  mutate(data_available = data_type)
}

``` 

```{r error=FALSE, warning=FALSE, message=FALSE}

# Transform all live births data
# Will abbreviate this to lapply(data[1:2], ur_transform) and lapply(data[3:6], ap_transform)
bt_ur_clean1 <- ur_transform(all_bts_excel[[1]], "births_urbanrural")
bt_ur_clean2 <- ur_transform(all_bts_excel[[2]], "births_urbanrural")
bt_am_clean1 <- ap_transform(all_bts_excel[[3]], "births_age_mother")
bt_am_clean2 <- ap_transform(all_bts_excel[[4]], "births_age_mother")
bt_af_clean1 <- ap_transform(all_bts_excel[[5]], "births_age_father")
bt_af_clean2 <- ap_transform(all_bts_excel[[6]], "births_age_father")
bt_ur_all <- join_transform(bt_ur_clean1, bt_ur_clean2)
bt_am_all <- join_transform(bt_am_clean1, bt_am_clean2)
bt_af_all <- join_transform(bt_af_clean1, bt_af_clean2)

```

```{r error=FALSE, warning=FALSE, message=FALSE}

# Create Freq Table Function for Supplementary Tables
st_clean <- function(data) {
    data %>% 
    clean_names() %>% 
    group_by(country_or_area, year) %>%
    summarise(n=n()) %>% 
    dplyr::select(country_or_area, year) %>% 
    rename("countries_areas" = country_or_area) %>% 
    group_by(countries_areas) %>% 
    summarise(n=n())
  }
```

```{r error=FALSE, warning=FALSE, message=FALSE}

# Load Supplementary Tables
all_lb <- list.files(path = "./Birth Regs", # insert path to folder
                        pattern = "LB", # do not change this
                        full.names = TRUE)

all_lbs_csv <-
  lapply(all_lb, function (i) {
  x = read_csv(i)
}) 

```


```{r error=FALSE, warning=FALSE, message=FALSE}

# Create Frequency Tables for All STs

all_births_long <-
  lapply(all_lbs_csv, function (i) {
  x = st_clean(i)
}) 

all_births_long <-
  Map(cbind, all_births_long, data_available = 
        list("birth_order", "birth_bw_sex","birth_gestation",  
             "birth_month", "birth_plural", "birth_ord_sex", 
             "birth_mar_len"))

all_births_long <- rbindlist(all_births_long)
all_births_long <- select(all_births_long, c(countries_areas, data_available, n))
all_births_tables <- bind_rows(all_births_long, bt_ur_all, bt_af_all, bt_am_all)
```

```{r error=FALSE, warning=FALSE, message=FALSE}

# Create Live Births Lookup
types_birth_long <- c("birth_order", "birth_bw_sex","birth_gestation",  "birth_month",
                      "birth_plural", "birth_ord_sex", "birth_mar_len", "births_urbanrural",
                      "births_age_mother", "births_age_father")
lookup_birth_unfpa <- lookup_score(types_birth_long, all_births_tables)

```

```{r error=FALSE, warning=FALSE, message=FALSE}

# Create UNFPA-specific disaggregated Live Births data
all_births_unfpa <- left_join(lookup_birth_unfpa, all_births_tables, by = c("countries_areas", "data_available"))

births_unfpa_clean <-
  all_births_unfpa %>% 
  mutate_all(~replace(., is.na(.), 0)) %>% 
  mutate(score = ((complete/5)*0.85)+((x/50)*0.15)) %>% 
  mutate (countries_areas = reorder(countries_areas, score))

```


```{r fig.height = 8, fig.width = 10}

# Generate Live Birth Heatmaps

birth_labels <- c("Urban & Rural", "Age of Mother*", "Age of Father*", "Order & Age of Mother", 
                  "Order & Sex", "Birth Month", "Weight & Sex", "Gestational Age", 
                  "Marriage Duration\nof Parents", "Birth Plurality")
birth_prefs <- c("births_urbanrural", "births_age_mother", "births_age_father", 
                 "birth_order", "birth_ord_sex", "birth_month", "birth_bw_sex", "birth_gestation", 
                 "birth_mar_len", "birth_plural")

det_heatmap(births_unfpa_clean, "AP", "Live Births", birth_labels, birth_prefs)
det_heatmap(births_unfpa_clean, "AS", "Live Births", birth_labels, birth_prefs)
det_heatmap(births_unfpa_clean, "EECA", "Live Births", birth_labels, birth_prefs)
det_heatmap(births_unfpa_clean, "ESA", "Live Births", birth_labels, birth_prefs)
det_heatmap(births_unfpa_clean, "LAC", "Live Births", birth_labels, birth_prefs)
det_heatmap(births_unfpa_clean, "WCA", "Live Births",birth_labels, birth_prefs)
```

## Marriages and Divorces

```{r error=FALSE, warning=FALSE, message=FALSE}

# Load Marriages and Divorces Tables

all_mdt <- list.files(path = "./MarriageDiv", 
                        pattern = "table",
                        full.names = TRUE)

all_mdt_excel <-
  lapply(all_mdt, function (i) {
  x = read_excel(i, skip = 4)
}) 

md_am_clean1 <- ap_transform(all_mdt_excel[[1]], "marriage_age")
md_am_clean2 <- ap_transform(all_mdt_excel[[2]], "marriage_age")
md_am_all <- join_transform(md_am_clean1, md_am_clean2)
```

```{r error=FALSE, warning=FALSE, message=FALSE}

# Load Supplementary Tables :Marriages and Divorces

all_mds <- list.files(path = "./MarriageDiv", 
                        pattern = "MD", 
                        full.names = TRUE)

all_mds_csv <-
  lapply(all_mds, function (i) {
  x = read_csv(i)
}) 


```

```{r error=FALSE, warning=FALSE, message=FALSE}

# Transform all Marriages and Divorces STs into Frequencies

all_mardivs_long <-
  lapply(all_mds_csv, function (i) {
  x = st_clean(i)
}) 

all_mardivs_long <-
  Map(cbind, all_mardivs_long, data_available = 
        list("marriage_cross", "divorce_ur", "marriage_1st_age", "marriage_ur"))

all_mardivs_long <- rbindlist(all_mardivs_long)
all_mardivs_long <- select(all_mardivs_long, c(countries_areas, data_available, n))
all_mardivs_tables <- bind_rows(all_mardivs_long, md_am_clean1, md_am_clean2)
```

```{r error=FALSE, warning=FALSE, message=FALSE}

# Create Marriage and Divorces Lookup
lookup_mardiv_unfpa <- lookup_score(c("marriage_cross", "divorce_ur", 
                                      "marriage_1st_age", "marriage_ur", "marriage_age"), 
                                    all_mardivs_tables)

```

```{r error=FALSE, warning=FALSE, message=FALSE}

# Generate all UNFPA-specific marriage and divorce availability data 
all_mardiv_unfpa <- left_join(lookup_mardiv_unfpa, all_mardivs_tables, by = c("countries_areas", "data_available"))

mardiv_unfpa_clean <-
  all_mardiv_unfpa %>% 
  mutate_all(~replace(., is.na(.), 0)) %>% 
  mutate(score = ((complete/5)*0.85)+((x/50)*0.15)) %>% 
  mutate (countries_areas = reorder(countries_areas, score))

```


```{r fig.height = 6.5, fig.width = 10}

# Generate Marriage/Divorces Heatmaps

mardiv_levels <- c("marriage_ur", "marriage_cross", "marriage_age", "marriage_1st_age", "divorce_ur")
mardiv_labels <- c("Urban/Rural\n(Marriages)", "Prior Marital Status\n(Marriages)", 
                   "Age\n(Marriages)", "Age at First Marriage\n(Marriages)", "Urban/Rural\n(Divorces)")

det_heatmap(mardiv_unfpa_clean, "AP", "Marriages and Divorces", mardiv_labels, mardiv_levels)
det_heatmap(mardiv_unfpa_clean, "AS", "Marriages and Divorces", mardiv_labels, mardiv_levels)
det_heatmap(mardiv_unfpa_clean, "EECA", "Marriages and Divorces", mardiv_labels, mardiv_levels)
det_heatmap(mardiv_unfpa_clean, "ESA", "Marriages and Divorces", mardiv_labels, mardiv_levels)
det_heatmap(mardiv_unfpa_clean, "LAC", "Marriages and Divorces",mardiv_labels, mardiv_levels)
det_heatmap(mardiv_unfpa_clean, "WCA", "Marriages and Divorces", mardiv_labels, mardiv_levels)

```

```{r error=FALSE, warning=FALSE, message=FALSE}

# Supplementary Tables :General and Foetal Deaths

all_gds <- list.files(path = "./Gen Death", 
                        pattern = "GD", 
                        full.names = TRUE)

all_gds_csv <-
  lapply(all_gds, function (i) {
  x = read_csv(i)
}) 

all_fds <- list.files(path = "./Gen Death", 
                        pattern = "FD", 
                        full.names = TRUE)

all_fds_csv <-
  lapply(all_fds, function (i) {
  x = read_csv(i)
}) 

id_agesex<-read_excel("./Gen Death/IFagesex.xlsx")
id_urbanrural<-read_csv("./Gen Death/IFurbanrural.csv")

```

```{r error=FALSE, warning=FALSE, message=FALSE}
all_gds_long <-
  lapply(all_gds_csv, function (i) {
  x = st_clean(i)
}) 

all_gds_long <-
  Map(cbind, all_gds_long, data_available = 
        list("gm_cause", "gm_age_sex_ur", "gm_month", "gm_sex_ur"))

all_fds_long <-
  lapply(all_fds_csv, function (i) {
  x = st_clean(i)
}) 

all_fds_long <-
  Map(cbind, all_fds_long, data_available = 
        list("ab_urbanrural", "fd_agewoman", "fd_gest_age", "fd_sex_ur"))

id_agesex_long<- ap_transform(id_agesex, "inf_death_age_sex")

id_urbanrural_long <-
  id_urbanrural %>% 
  st_clean() %>% 
  cbind(data_available = "inf_death_urbanrural")


all_gfd_tables <- bind_rows(rbindlist(all_gds_long), 
                            rbindlist(all_fds_long), all_fds_long, id_agesex_long, 
                            id_urbanrural_long)
```

```{r error=FALSE, warning=FALSE, message=FALSE}

# I just identified a problem in the wrangling of Infant Death tabs by age and sex. I will come back to this.
id_agesex_long<-
  id_agesex[- 1, ] %>%
  select (1:2) %>% 
  rename_all(~c("countries_areas", "total"))   
```

```{r error=FALSE, warning=FALSE, message=FALSE}
lookup_gfd_unfpa <- lookup_score(c("inf_death_urbanrural", 
                                   "gm_cause", "gm_age_sex_ur", "gm_month", "gm_sex_ur", 
                                   "ab_urbanrural", "fd_agewoman", "fd_gest_age", "fd_sex_ur"), 
                                 all_gfd_tables)
```

```{r error=FALSE, warning=FALSE, message=FALSE}
all_gfd_unfpa <- left_join(lookup_gfd_unfpa, all_gfd_tables, by = c("countries_areas", "data_available"))
```

```{r error=FALSE, warning=FALSE, message=FALSE}
gfd_unfpa_clean <-
  all_gfd_unfpa %>% 
  mutate_all(~replace(., is.na(.), 0)) %>% 
  mutate(score = ((complete/5)*0.85)+((x/50)*0.15)) %>% 
  mutate (countries_areas = reorder(countries_areas, score))
```


```{r fig.height = 6.5, fig.width = 10}

GF_levels <- c("fd_sex_ur", "fd_agewoman", "fd_gest_age", "inf_death_urbanrural", "ab_urbanrural", "gm_sex_ur", "gm_age_sex_ur", "gm_cause", "gm_month")
GF_labels <- c("Urban/Rural\n(Foetal Deaths)", "Age of Mother\n(Foetal Deaths)", "Gestational Age\n(Foetal Deaths)", "Urban/Rural\n(Infant Deaths)", "Age & Sex\n(Infant Deaths)", "Urban/rural\n(Abortions)", "Sex & Urban/Rural\n(General Deaths)", "Age, Sex, & Urban/Rural\n(General Deaths)", "Cause\n(General Deaths)", "Month\n(General Deaths)")

det_heatmap(gfd_unfpa_clean, "AP", "Foetal Deaths, Abortions, and General Deaths\n", GF_labels, GF_levels)
det_heatmap(gfd_unfpa_clean, "AS", "Foetal Deaths, Abortions, and General Deaths\n", GF_labels, GF_levels)
det_heatmap(gfd_unfpa_clean, "EECA", "Foetal Deaths, Abortions, and General Deaths\n", GF_labels, GF_levels)
det_heatmap(gfd_unfpa_clean, "ESA", "Foetal Deaths, Abortions, and General Deaths\n", GF_labels, GF_levels)
det_heatmap(gfd_unfpa_clean, "LAC", "Foetal Deaths, Abortions, and General Deaths\n", GF_labels, GF_levels)
det_heatmap(gfd_unfpa_clean, "WCA", "Foetal Deaths, Abortions, and General Deaths\n", GF_labels, GF_levels) 
# Will fix this title wrapping


```


# Create Steps to Include Complete Data of Tables with "latest available year"

Live Births
-Table 10 - Live births by age of mother and sex of child, general and age-specific fertility rates: latest available year, 2009-2018
-Table 11 - Live births and live birth rates by age of father: latest available year, 2009-2018

Marriages
-Table 23 - Marriages by age of groom and by age of bride: latest available year, 2009-2018

Divorces
-Table 25 - Divorces and percentage distribution by duration of marriage, latest available year: 2009-2018

Foetal Deaths and Abortions
-Table 14 - Legally induced abortions by age and number of previous live births of women: latest available year, 2009-2018

Infant Deaths
-Table 16 - Infant deaths and infant mortality rates by age and sex, latest available year: 2009-2018

General Deaths
-Table 19 - Deaths by age and sex, age-specific death rates by sex: latest available year, 2009-2018

```{r error=FALSE, warning=FALSE, message=FALSE}

# Test steps with Table 19 - General Death, Age and Sex
all_gds_age_sex <- list.files(path = "./Gen Death", # Load all General Deaths by Age and Sex, Latest Available Year
                        pattern = "table19", 
                        full.names = TRUE)
all_gds_age_sex_excel <- # Convert into Excel
  lapply(all_gds_age_sex, function (i) {
  x = read_excel(i)
}) 

```

```{r error=FALSE, warning=FALSE, message=FALSE}

# Create Freq Table Function for non-Urban/Rural DYB Tables, modified
la_transform_mod <- function(data_list, data_type) {
  
  data_list_tfr<-
  lapply(data_list, function (i) {
  x = ap_transform(i, data_type)
}) 
    
  data_1<-rbindlist( data_list_tfr)
  
   data_1<-
    data_1 %>% 
    group_by(countries_areas) %>% 
    summarise(n=n()) %>% 
    mutate("data_available" = data_type)
   
  return(data_1)
}

```

```{r}

#Create dataset of previous years
all_gds_age_sex_long <- la_transform_mod(all_gds_age_sex_excel, "gd_age_sex")

```

## Test Visualization
```{r error=FALSE, warning=FALSE, message=FALSE}
lookup_gfd_unfpa_mod <- lookup_score(c("inf_death_age_sex", "inf_death_urbanrural", 
                                       "gm_cause", "gm_age_sex_ur", "gm_month", "gm_sex_ur", 
                                       "gd_age_sex", "ab_urbanrural", "fd_agewoman", "fd_gest_age", 
                                       "fd_sex_ur"), all_gfd_tables)
```

```{r error=FALSE, warning=FALSE, message=FALSE}
all_gfd_tables_mod <- bind_rows(rbindlist(all_gds_long), rbindlist(all_fds_long), 
                                all_fds_long, id_agesex_long, all_gds_age_sex_long, id_agesex_long, id_urbanrural_long)
all_gfd_unfpa_mod <- left_join(lookup_gfd_unfpa_mod, all_gfd_tables_mod, 
                               by = c("countries_areas", "data_available"))
```

```{r error=FALSE, warning=FALSE, message=FALSE}
gfd_unfpa_clean_mod <-
  all_gfd_unfpa_mod %>% 
  mutate_all(~replace(., is.na(.), 0)) %>% 
  mutate(score = ((complete/5)*0.85)+((x/50)*0.15)) %>% 
  mutate (countries_areas = reorder(countries_areas, score))
```

```{r fig.height = 8, fig.width = 10}


GF_levels_mod <- c("fd_sex_ur", "fd_agewoman", "fd_gest_age", "inf_death_urbanrural", 
                   "inf_death_age_sex","ab_urbanrural", "gm_sex_ur", "gd_age_sex", 
                   "gm_age_sex_ur", "gm_cause", "gm_month")
GF_labels_mod <- c("Urban/Rural\n(Foetal Deaths)", "Age of Mother\n(Foetal Deaths)", 
                   "Gestational Age\n(Foetal Deaths)", "Urban/Rural\n(Infant Deaths)", 
                   "Age & Sex\n(Infant Deaths)", "Urban/rural\n(Abortions)", "Sex & Urban/Rural\n(General Deaths)", 
                   "Age, Sex, & Urban/Rural\n(General Deaths)", "Age & Sex\n(General Deaths)", "Cause\n(General Deaths)", "Month\n(General Deaths)")

det_heatmap(gfd_unfpa_clean_mod, "AP", "Foetal Deaths, Abortions, and General Deaths\n", GF_labels_mod, GF_levels_mod)
det_heatmap(gfd_unfpa_clean_mod, "AS", "Foetal Deaths, Abortions, and General Deaths\n", GF_labels_mod,  GF_levels_mod)
det_heatmap(gfd_unfpa_clean_mod, "EECA", "Foetal Deaths, Abortions, and General Deaths\n", GF_labels_mod,  GF_levels_mod)
det_heatmap(gfd_unfpa_clean_mod, "ESA", "Foetal Deaths, Abortions, and General Deaths\n", GF_labels_mod,  GF_levels_mod)
det_heatmap(gfd_unfpa_clean_mod, "LAC", "Foetal Deaths, Abortions, and General Deaths\n", GF_labels_mod,  GF_levels_mod)
det_heatmap(gfd_unfpa_clean_mod, "WCA", "Foetal Deaths, Abortions, and General Deaths\n",GF_labels_mod,  GF_levels_mod)


```


